---
title: "Workflow Cost Report `r Sys.Date()`"
data: "`r Sys.Date()`"
output: flexdashboard::flex_dashboard
params:
  workspace: !r unique(list.files("../data"))[1]
  sampleType: !r read.table(paste0("../data/", unique(list.files("../data"))[1], "/workflowData.tsv"), header = T, sep='\t')$entityType[1]
---

```{r global, include=FALSE}
library(data.table)
library(htmltools)
library(plotly)
library(shiny)
library(DT)
library(stringr)
library(ComplexHeatmap)
library(InteractiveComplexHeatmap)
library(circlize)
library(ggplot2)
library(patchwork)


```


```{r loading_data}
mainTbl <- fread(paste0("../data/", params$workspace, "/workflowData.tsv"))[cost != "N/A",
                                                                            .(workflow, workflowId, sample, entityType, cost = as.numeric(cost), costType, status,
                                                                             statusLastChangedDate, workflowVersion)]

storData <- fread(paste0("../data/", params$workspace, "/StorageEstimate.tsv"))

charrTbl <- data.table()
charrPath <- paste0("../data/", params$workspace, "/", params$sampleType, "_charr.tsv")
if (file.exists(charrPath)) {
  charrTbl <- fread(charrPath)
}

demoTbl <- data.table()
demoPath <- paste0("../data/", params$workspace, "/", params$sampleType, "_demo.tsv")
if (file.exists(demoPath)) {
  demoTbl <- fread(demoPath)
}
```

# Current Sample Status

## Column 1 {data-width="300"}

### Sample Status Heatmap {data-height="600"}

```{r progress_heatmap, fig.width = 6, fig.height = 7}
# color palette for progress heatmap, no need to change
pal <- colorRampPalette(c("skyblue", "gray", "red", "green"))(4)
colors = structure(pal, names = c("0", "1", "2", "3"))
valmap = c("Running" = 0, "Not Started" = 1, "Failed" = 2, "Aborted" = 2, "Succeeded" = 3)


mapTable <- dcast(
  unique(mainTbl[entityType == params$sampleType][order(status, decreasing = TRUE)], by = c("workflow", "sample"))[, c("workflow", "sample", "status")],
  sample ~ workflow,
  value.var = "status"
)
wfCols <- colnames(mapTable)[-1]
mapTable[, (wfCols) := lapply(.SD, function(col) fifelse(is.na(col), "Not Started", col)), .SDcols = wfCols]

mapTable$progress <- rowSums(mapTable[,-1] == "Succeeded")
setorderv(mapTable, c("progress", wfCols), order = -1)

# logic for coloring table
mapTable[, (wfCols) := lapply(
  .SD,
  function(col) valmap[col]
), .SDcols = wfCols]


Heatmap(
  mapTable[, wfCols, with = FALSE],
        width = unit(8, "cm"),
        height = unit(12, "cm"),
        col = colors,
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        column_names_side = "top",
        column_names_rot = 45,
        column_names_gp = gpar(fontsize = 10, fontface = "bold"),
        rect_gp = gpar(col = "darkgray", lwd = 0.1),
        border_gp = gpar(col = "black", lwd = 0.2),
        heatmap_legend_param = list(
          title = "Status",
          at = 0:3,
          color_bar = "discrete",
          labels = c("Running", "Not Started", "Failed", "Complete"),
          labels_gp = gpar(fontsize = 10),
          title_gp = gpar(fontsize = 12, fontface = "bold"),
          grid_height = unit(8, "mm"),
          grid_width = unit(8, "mm")
        ),
        left_annotation = rowAnnotation(
          title = anno_block(
            labels = c(paste0(nrow(mapTable), "\nTotal Samples")),
            labels_rot = 0,
            width = unit(3, "cm"),
            gp = gpar(col = NA)
          )
        )
)


```

### Cost Distribution {data-height="400"}

```{r, fig.height = 3, fig.width = 5}

tmpTbl <- mainTbl[entityType == params$sampleType][, .(total_cost = round(sum(as.numeric(cost)), 2)), by = c("workflow", "sample")]
tmpTbl <- tmpTbl[, .(total_cost = sum(total_cost)), by = sample]

sampStatus <- unique(mainTbl[entityType == params$sampleType &
                               status == "Succeeded", c("workflow", "sample")])
numWorkflows <- uniqueN(sampStatus$workflow)
sampStatus <- sampStatus[,.(count = .N), by = sample]
completeSamples <- sampStatus[count == numWorkflows]$sample


ggplot(rbind(tmpTbl[, .(type = "All"), c("total_cost")],
             tmpTbl[sample %in% completeSamples, .(type = "Complete"), c("total_cost")]),
       aes(total_cost, type)) +
  geom_boxplot() +
  ylab("") +
  xlab("Total Cost per Sample ($)") +
  scale_x_continuous(limits = c(0, 15)) +
  ggtitle("Sample Costs") +
  theme_linedraw() +
  theme(text = element_text(size=14))

```

## Column 2 {data-width="700"}

### Sample Status Table

```{r}

statTable <- unique(mainTbl[entityType == params$sampleType][,c("sample", "workflow", "status")][order(-status)], by = c("sample", "workflow"))
datatable(dcast(statTable[,.(count=.N),by=c("workflow","status")], workflow ~ status, fill = 0))

```

### Sample Cost Table

```{r}

tmpCostTbl <- mainTbl[entityType == params$sampleType][, .(total_cost = round(sum(as.numeric(cost)), 2)), by = c("workflow", "sample")]


sampByWf <- dcast(tmpCostTbl, sample ~ workflow, value.var = "total_cost")
sampByWf[, total := round(rowSums(.SD, na.rm = TRUE), 2), .SDcols = colnames(sampByWf)[-1]]
datatable(sampByWf)

```

# Workflow Cost Estimates

## Column 1 {data-width="350"}

### Workflow Distribution

```{r workflow_dist, echo=FALSE}

tmpCostTbl <- mainTbl[entityType == params$sampleType][, .(total_cost = round(sum(as.numeric(cost)), 2)), by = c("workflow", "sample")]

ggplot(tmpCostTbl, aes(total_cost, workflow)) +
  geom_boxplot() +
  ylab("") +
  scale_x_continuous(limits = c(0, 10)) +
  ggtitle("Workflow Costs per Sample ($)") +
  theme_linedraw() +
  theme(text = element_text(size = 14))

```

### Workflow Cost Table

```{r}

tmpCostTbl <- mainTbl[entityType == params$sampleType][, .(total_cost = round(sum(as.numeric(cost)), 2)), by = c("workflow", "sample")]
datatable(tmpCostTbl[, .(avg_cost = mean(total_cost), sd_cost = sd(total_cost),
                       total = sum(total_cost)), by = workflow]) %>% formatRound(c(2:4), 2)


```

## Column 2 {data-width="650"}

### All Workflows

```{r whole_table, echo=FALSE}

datatable(mainTbl[entityType == params$sampleType])
```

### Workflow Table

```{r workflow_table, echo=FALSE}

tmpCost <-   mainTbl[entityType == params$sampleType][, .(total_cost = sum(cost)), by = c("workflow", "workflowVersion", "sample")][
  ,.(avg_cost = mean(total_cost), sd_cost = sd(total_cost), count = .N, total_cost = sum(total_cost)),
  by = c("workflow", "workflowVersion")][order(workflow, workflowVersion)
]
datatable(tmpCost) %>% formatRound(c(3,4,6), 2)

```

# Data Usage

## Column 1

### Storage Cost Estimate

```{r}
sizeFactor <- 1024 ^ 4 # TB
sizeUnit <- "TB"


liveCost <- round(storData$`usage.live-object`[1] / sizeFactor, 2)
softCost <- round(storData$`usage.soft-deleted-object`[1] / sizeFactor, 2)
if (liveCost + softCost > 0) {
  pie(c(liveCost, softCost),
    labels = c(paste0(liveCost, sizeUnit, " (Live)"),
               paste0(softCost, sizeUnit, " (Soft-delete)")),
    main = paste0("Storage Cost (Monthly Total = $", storData$estimate, ")"),
    cex = 1,
    cex.main = 1.2)
}


```

# QC Stats

## Column 1 {data-width=600}

### Charr QC Table

```{r}
datatable(charrTbl)
```

## Column 2 {data-width=400}

### Charr QC Plots

```{r, fig.height = 6, fig.width = 6}
if (nrow(charrTbl) > 0) {
  charrPlts <- lapply(colnames(charrTbl)[-1], function(col) {
    ggplot(charrTbl[, c("#SAMPLE", col), with = F], aes(.data[[col]])) + geom_boxplot() + theme_classic() +
      scale_x_continuous(labels = function(x) {
        ifelse(x > 1000, scales::label_scientific()(x), x)
      }) +
      xlab("") + 
      theme(axis.ticks.y = element_blank(), axis.text.y = element_blank()) +
      ggtitle(col)
  })
  wrap_plots(charrPlts, nrow = 4)
}


```

# Demographics

## Column 1

### Sex and Ancestry Table

```{r}
datatable(demoTbl)
```

## Column 2

### Ancestry Plot {data-height=600}

```{r, fig.width = 8, fig.height = 3}
if (nrow(demoTbl) > 0){
  anc_colors <- setNames(
    scales::hue_pal()(length(unique(demoTbl$ancestry))),
    unique(demoTbl$ancestry)
  )
  
  ancBar <- ggplot(demoTbl, aes(ancestry, fill = ancestry)) +
    geom_bar() +
    xlab("") + 
    scale_fill_manual(values = anc_colors) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    theme_classic() +
    theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
  
  ancPts <- ggplot(demoTbl, aes(`GD1(x)`, `GD2(y)`, color = ancestry)) +
    geom_point() +
    scale_color_manual(values = anc_colors) +
    theme_classic() +
    theme(legend.position = "none")
  
  ancPts + ancBar
}

```

### Sex Karyotype Counts {data-height=400}

```{r}
if (nrow(demoTbl) > 0) {
  datatable(demoTbl[,.(count = .N), by = c("chrX_count", "chrY_count")])
}
```