---
title: "Workflow Cost Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```

```{r packages, include=FALSE}
packages <- c("ggplot2", "circlize", "data.table", "stringr", "ComplexHeatmap")
install.packages(setdiff(packages, rownames(installed.packages())))  

library(shiny)
library(data.table)
library(DT)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
```

```{r load_data, include=FALSE}
runData <- fread("~/Desktop/Projects/CostMonitor/data/workflowData.tsv")
batchData <- runData[entityType == "sample_set" & !is.na(cost)]
runData <- runData[entityType == "sample"]
runData <- runData[, c("workflow", "workflowId", "sample", "cost", "costType", "status", "statusLastChangedDate", "workflowVersion")]
runData <- runData[cost != "N/A"]
runData[, cost := as.numeric(cost)]
```

```{r prep_data, include=FALSE}
# cost per sample/workflow
costSampWf <- runData[, .(total_cost = round(sum(as.numeric(cost)), 2)), by = c("workflow", "sample")]
costSamp <- costSampWf[, .(total_cost = sum(total_cost)), by = sample]
costWf <- runData[, .(total_cost = sum(cost)), by = c("workflow", "workflowVersion", "sample")][, .(avg_cost = mean(total_cost), sd_cost = sd(total_cost), count = .N, total_cost = sum(total_cost)), by = c("workflow", "workflowVersion")][order(workflow, workflowVersion)]

statusTable <- dcast(
  unique(runData[order(status, decreasing = TRUE)], 
         by = c("workflow", "sample"))[, c("workflow", "sample", "status")][, status := as.integer(status == "Succeeded")],
  sample ~ workflow,
  value.var = "status"
)

wfCols <- colnames(statusTable)[-1]
statusTable$progress <- rowSums(statusTable[,-1], na.rm = TRUE)
setorderv(statusTable, c("progress", wfCols), order = -1)

invStatus <- copy(statusTable)
invStatus[, (wfCols) := lapply(.SD, function(col) as.integer(col == 0)), .SDcols = wfCols]
statusSumm <- rbind(colSums(statusTable[, -c("sample", "progress")], na.rm = TRUE),
                    colSums(invStatus[, -c("sample", "progress")], na.rm = TRUE))
rownames(statusSumm) <- c("Success", "Fail")

completeSamples <- statusTable[progress == length(wfCols)]$sample 
fullAvg <- mean(costSamp[sample %in% completeSamples]$total_cost)
fullSD <- sd(costSamp[sample %in% completeSamples]$total_cost)
```

# Current Sample Status

## Column {.sidebar}

```{r progress_heatmap_inputs}
checkboxGroupInput("wfCheck", "Workflows to show", 
                   choices = colnames(statusTable[,-c("sample", "progress")]),
                   selected = colnames(statusTable[,-c("sample", "progress")]))
```

## Column 1 {data-width=300}

### Sample Status Heatmap {data-height=500}

```{r progress_heatmap, fig.height = 9, fig.width = 7}
mapTable <- statusTable
mapTable[, (wfCols) := lapply(.SD, function(col) fifelse(is.na(col), 1, col + 2)), .SDcols = wfCols]

pal <- colorRampPalette(c("skyblue", "gray", "red", "green"))(4)
colors = structure(pal, names = c("0", "1", "2", "3"))
setorderv(mapTable, wfCols, order = -1)

renderPlot({
  hmap <- Heatmap(mapTable[,input$wfCheck, with = FALSE], 
          width = unit(15, "mm") * ncol(statusTable),
          height = unit(12, "cm"),
          col = colors,
          cluster_columns = FALSE,
          cluster_rows = FALSE,
          column_names_side = "top",
          column_names_rot = 45,
          column_names_gp = gpar(fontsize = 10, fontface = "bold"),
          rect_gp = gpar(col = "darkgray", lwd = 0.1),
          border_gp = gpar(col = "black", lwd = 0.2),
          heatmap_legend_param = list(
            title = "Status",
            at = 0:3,
            color_bar = "discrete",
            labels = c("G2C processed", "not started", "failed", "complete"),
            labels_gp = gpar(fontsize = 10),
            title_gp = gpar(fontsize = 12, fontface = "bold"),
            grid_height = unit(8, "mm"),
            grid_width = unit(8, "mm")
          ),
          left_annotation = rowAnnotation(
            title = anno_block(
              labels = c(paste0(nrow(mapTable), "\nTotal Samples")), 
              labels_rot = 0, 
              width = unit(3, "cm"), 
              gp = gpar(col = NA)
            )
          )
  )
  draw(hmap, padding = unit(c(0,0,0,0), "mm")) 
})

```

### Cost Distribution {data-height=200}

```{r cost_distribution, echo=FALSE, fig.width = 5, fig.height = 3}
renderPlot({
  costSamp <- costSampWf[workflow %in% input$wfCheck][, .(total_cost = sum(total_cost)), by = sample]
  ggplot(rbind(costSamp[, .(type = "All"), c("total_cost")], 
               costSamp[sample %in% completeSamples, .(type = "Complete"), c("total_cost")]),
         aes(total_cost, type)) + 
    geom_boxplot() +
    ylab("") +
    xlab("Total Cost per Sample ($)") +
    scale_x_continuous(limits = c(0, 15)) +
    ggtitle("Sample Costs") +
    theme_linedraw() +
    theme(text = element_text(size=20))
})

```

## Column 2

### Sample Status Table {data-width=100, data-height=400}

```{r sample_status_table, echo=FALSE}
renderDataTable({
  datatable(t(statusSumm[, input$wfCheck]))
})
```

### Sample Cost Table {data-height=600}

```{r sample_cost_table, echo=FALSE}
sampByWf <- dcast(costSampWf, sample ~ workflow, value.var = "total_cost")
sampByWf[, total := round(rowSums(.SD), 2), .SDcols = colnames(sampByWf)[-1]]
renderDataTable({
  sampByWf[,c("sample", input$wfCheck, "total"), with = FALSE]
})
```

# Workflow Cost Estimates

## Column 1 {data-width=300}

### Workflow Distribution {data-height=500}

```{r workflow_dist, echo=FALSE, fig.width = 5, fig.height = 3}
renderPlot({
  ggplot(costSampWf, aes(total_cost, workflow)) +
    geom_boxplot() +
    ylab("") +
    scale_x_continuous(limits = c(0, 10)) +
    ggtitle("Workflow Costs per Sample ($)") +
    theme_linedraw()
})

```

### Workflow Cost Table {data-height=500}
 
```{r workflow_cost_table, echo=FALSE}
datatable(costSampWf[, .(avg_cost = mean(total_cost), sd_cost = sd(total_cost),
                         total = sum(total_cost)), by = workflow]) %>% formatRound(c(2:4), 2)

```
 
## Column 2 {data-width=500}

### All Workflows {data-height=500}

```{r whole_table, echo=FALSE}
datatable(runData[,-c(2,7)])
```

### Workflow Table {data-height=500}

```{r workflow_table, echo=FALSE}
datatable(costWf) %>% formatRound(c(3,4,6), 2)
```

# Storage Cost Estimates

## Column 1

### Storage Cost Chart

```{r storage, echo=FALSE, fig.height=10, fig.width=12}
storData <- fread("../data/StorageEstimate.tsv")
sizeFactor <- 1024 ^ 4 # TB
sizeUnit <- "TB"
liveCost <- round(storData$`usage.live-object`[1] / sizeFactor, 2)
softCost <- round(storData$`usage.soft-deleted-object`[1] / sizeFactor, 2)
pie(c(liveCost, softCost),
    labels = c(paste0(liveCost, sizeUnit, " (Live Objects)"), 
               paste0(softCost, sizeUnit, " (Soft-delete Objects)")),
    main = paste0("Storage Cost (Monthly Total = $", storData$estimate, ")"))
```

# Data Usage

## Column 1

### Per-Workflow Use

```{r log_summary, echo=FALSE}
logData <- fread("../data/logSummary.tsv")

### Plan to fix the formatting in the generating script ###
newCols <- colnames(logData)[-1]
logData[, avg_write := NULL]
colnames(logData) <- newCols
### Should not need this ###

# remove min and max for now
logData <- logData[, colnames(logData)[!(colnames(logData) %like% "min_|max_")], with = FALSE]

datatable(logData[, -c(1,2,4,6)][, lapply(.SD, function(x) round(mean(x), 2)), by = c("workflow_name", "task_name")])
```