---
title: "Workflow Cost Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
```

```{r packages, include=FALSE}
packages <- c("ggplot2", "circlize", "data.table", "stringr", "ComplexHeatmap")
install.packages(setdiff(packages, rownames(installed.packages())))  

library(data.table)
library(DT)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
```

```{r load_data, include=FALSE}
runData <- fread("~/Desktop/Projects/CostMonitor/data/workflowData.tsv")
runData <- runData[entityType == "sample"]
runData <- runData[, c("workflow", "workflowId", "sample", "cost", "costType", "status", "statusLastChangedDate")]
runData <- runData[cost != "N/A"]
runData[, cost := as.numeric(cost)]
```

```{r prep_data, include=FALSE}
# cost per sample/workflow
costSampWf <- runData[, .(total_cost = sum(as.numeric(cost))), by = c("workflow", "sample")]
costSamp <- costSampWf[, .(total_cost = sum(total_cost)), by = sample]
costWf <- costSampWf[, .(avg_cost = mean(total_cost), sd_cost = sd(total_cost), total_cost = sum(total_cost)), by = workflow]

statusTable <- dcast(
  unique(runData[order(status, decreasing = TRUE)], 
         by = c("workflow", "sample"))[, c("workflow", "sample", "status")][, status := as.integer(status == "Succeeded")],
  sample ~ workflow,
  value.var = "status"
)

wfCols <- colnames(statusTable)[-1]
statusTable$progress <- rowSums(statusTable[,-1], na.rm = TRUE)
setorderv(statusTable, c("progress", wfCols), order = -1)

invStatus <- copy(statusTable)
invStatus[, (wfCols) := lapply(.SD, function(col) as.integer(col == 0)), .SDcols = wfCols]
statusSumm <- rbind(colSums(statusTable[, -c("sample", "progress")], na.rm = TRUE),
                    colSums(invStatus[, -c("sample", "progress")], na.rm = TRUE))
rownames(statusSumm) <- c("Success", "Fail")

completeSamples <- statusTable[progress == length(wfCols)]$sample 
fullAvg <- mean(costSamp[sample %in% completeSamples]$total_cost)
fullSD <- sd(costSamp[sample %in% completeSamples]$total_cost)
```

# Current Sample Status

## Column 1 {data-width=400}

### Sample Status Heatmap {data-height=500}

```{r progress_heatmap, fig.height = 9, fig.width = 7}
mapTable <- statusTable
mapTable[, (wfCols) := lapply(.SD, function(col) fifelse(is.na(col), 1, col + 2)), .SDcols = wfCols]

pal <- colorRampPalette(c("skyblue", "gray", "red", "green"))(4)
colors = structure(pal, names = c("0", "1", "2", "3"))
setorderv(mapTable, wfCols, order = -1)

hmap <- Heatmap(mapTable[,-c("sample", "progress")], 
        width = unit(15, "mm") * ncol(statusTable),
        height = unit(12, "cm"),
        col = colors,
        cluster_columns = FALSE,
        cluster_rows = FALSE,
        column_names_side = "top",
        column_names_rot = 45,
        column_names_gp = gpar(fontsize = 10, fontface = "bold"),
        rect_gp = gpar(col = "darkgray", lwd = 0.1),
        border_gp = gpar(col = "black", lwd = 0.2),
        heatmap_legend_param = list(
          title = "Status",
          at = 0:3,
          color_bar = "discrete",
          labels = c("G2C processed", "not started", "failed", "complete"),
          labels_gp = gpar(fontsize = 10),
          title_gp = gpar(fontsize = 12, fontface = "bold"),
          grid_height = unit(8, "mm"),
          grid_width = unit(8, "mm")
        ),
        left_annotation = rowAnnotation(
          title = anno_block(
            labels = c(paste0(nrow(mapTable), "\nTotal Samples")), 
            labels_rot = 0, 
            width = unit(3, "cm"), 
            gp = gpar(col = NA)
          )
        )
)

draw(hmap, padding = unit(c(0,0,0,0), "mm"))
```

## Column 2

### Sample Status Table {data-width=100}

```{r sample_status_table, echo=FALSE}
datatable(t(statusSumm))
```

# Cost Estimates

## Column 1 {data-width=300}

### Cost Distribution

```{r cost_distribution, echo=FALSE, fig.width = 5, fig.height = 3}
ggplot(rbind(costSamp[, .(type = "All"), c("total_cost")], 
             costSamp[sample %in% completeSamples, .(type = "Complete"), c("total_cost")]),
       aes(total_cost, type)) + 
  geom_boxplot() +
  ylab("") +
  scale_x_continuous(limits = c(0, 15)) +
  ggtitle("Sample Costs ($)") +
  theme_linedraw()
```

### Workflow Distribution

```{r workflow_dist, echo=FALSE, fig.width = 5, fig.height = 3}
ggplot(costSampWf, aes(total_cost, workflow)) +
  geom_boxplot() +
  ylab("") +
  scale_x_continuous(limits = c(0, 10)) +
  ggtitle("Workflow Costs per Sample ($)") +
  theme_linedraw()
```
 
## Column 2 {data-width=500}

### All Workflows {data-height=600}

```{r whole_table, echo=FALSE}
datatable(runData[,-c(2,7)])
```

### Workflow Table {data-height=350}

```{r workflow_table, echo=FALSE}
datatable(costWf[, let(avg_cost = round(avg_cost, digits = 2), 
                       sd_cost = round(sd_cost, digits = 2),
                       total_cost = round(total_cost, digits = 2))])
```

# Workflow Fails

## Column 1

### Workflow Fail Distribution

```{r workflow_fail_dist, echo=FALSE}
costSampWfFail <- runData[status == "Failed", c("workflow", "sample", "cost")]
costSampWfFail$cost <- as.integer(ifelse(costSampWfFail$cost == "N/A", 0, costSampWfFail$cost))
ggplot(costSampWfFail, aes(cost, workflow)) +
  geom_boxplot() +
  ylab("") +
  scale_x_continuous(limits = c(0, 10)) +
  ggtitle("Distribution of Workflow Failure Costs per Run ($)") +
  theme_linedraw()
```

## Column 2

### Workflow Fail Table

```{r workflow_fail_table, echo=FALSE}
datatable(costSampWfFail[, .(avg_cost = round(mean(cost), 2), 
                                sd_cost = round(sd(cost), 2), 
                                total_cost = sum(cost)), by = "workflow"])
```

# Storage Cost Estimates

## Column 1

### Storage Cost Chart

```{r storage, echo=FALSE}
storData <- fread("../data/StorageEstimate.tsv")
sizeFactor <- 1024 ^ 4 # TB
sizeUnit <- "TB"
liveCost <- round(storData$`usage.live-object`[1] / sizeFactor, 2)
softCost <- round(storData$`usage.soft-deleted-object`[1] / sizeFactor, 2)
pie(c(liveCost, softCost),
    labels = c(paste0(liveCost, sizeUnit, " (Live Objects)"), 
               paste0(softCost, sizeUnit, " (Soft-delete Objects)")),
    main = paste0("Storage Cost (Monthly Total = $", storData$estimate, ")"))
```
